---
output: 
  github_document
--- 

# Area Under the Repeated Measures Curve

Zachary R. McCaw <br>
Updated: 2022-08-20

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

## Description

This package performs inference on the area under the repeated measures curve.

## Installation

```{R, eval=FALSE}
devtools::install_github(repo = 'zrmacc/AURMC')
```

## Data

The function `GenData` simulates example data in the format expected by this package. 

```{R}
library(AURMC)
set.seed(101)
data <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4
)
head(data)
```

The columns are:

* `idx`, the subject identifier. 
* `time`, the observation time. 
* `status`, coded 0 for censoring, 1 for a measurement, and 2 for a terminal event (e.g. death).
* `value`, the value of the measurement at the observation time.

Note that:

* All subjects must have a baseline record at `time = 0` with `status = 1`.
* All subjects must have an **observation-terminating** event as their last record, either `status = 0` for censoring or `status = 2` for a terminal event.
  - By default, a subject with no observation-terminating event is censored immediately after their last measurement.
* If a subject's measurement value is missing (`NA`), their last value is carried forward.
  - If a subject's baseline measurement is missing, it is set to zero.

## Interolation

The function `InterpolateR` provides the ability to linearly interpolate measurements to obtain more precise area estimates:
```{r}
# Simple example data.
data <- data.frame(
  idx = c(1, 1, 2, 2),
  status = c(1, 0, 1, 2),
  time = c(0, 1, 0, 1),
  value = c(0, 1, 0, -1)
)
show(data)
```

```{r}
grid <- c(0, 0.5, 1.0)
interpolated <- AURMC::Interpolate(
  data = data,
  grid = grid
)
show(interpolated)
```

## One-sample Problem

Case where the expected area is zero:
```{r}
set.seed(101)
data <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4
)
AURMC::AURMC(data, tau = 2.0)
```

Case where the expected area is non-zero:
```{r}
set.seed(101)
data <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4,
  value_mean = 1.0
)
AURMC::AURMC(data, tau = 2.0)
```

## Two-sample Problem

Case of no true difference:
```{r}
set.seed(102)

# Reference arm.
data0 <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4,
  value_mean = 1.0
)
data0$arm <- 0

# Treatment arm.
data1 <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4,
  value_mean = 1.0
)
data1$arm <- 1
data1$idx <- nrow(data0) + data1$idx

# Overall data set.
data <- rbind(data0, data1)

AURMC::CompareAURMCs(data)
```

Case of a true differnece:
```{r}
set.seed(102)

# Reference arm.
data0 <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4,
  value_mean = 1.0
)
data0$arm <- 0

# Treatment arm.
data1 <- AURMC::GenData(
  censoring_rate = 0.5,
  death_rate = 0.5,
  n = 25,
  tau = 4,
  value_mean = 2.0
)
data1$arm <- 1
data1$idx <- nrow(data0) + data1$idx

# Overall data set.
data <- rbind(data0, data1)

AURMC::CompareAURMCs(data)
```
